### 2020-07-30 theory/notes_banks_buyingfinitemajorities.tex
finish redacting notes for banks

Two states of the world: policies $x$ and $y$. Party $A$ and $B$ prefer $x$ and $y$ respectively. We can make this more explicit with
\begin{align*}
    U_A(x) > U_A(y)\\
    U_B(y) > U_B(x)
\end{align*}
For voters, $v_i > 0$ means voter $i$ prefers $x$ to $y$ All $i \in N$ vote for $x$ or $y$. Simple majority rule determines which policy gets implemented. For each voter, each party sets a bribe schedule
\begin{align*}
    a \in (a_1, ..., a_n) \in \mathbb{R}^n_+\\
    b \in (b_1, ..., b_n) \in \mathbb{R}^n_+
\end{align*}
Solving through backward induction, given bribe schedules $(a,b)$, voter $i$ prefers to vote for $x$ if $a_i + v_i > b_i$ and for $y$ otherwise. Since indifferent voters choose $y$, party $B$ needs to only match bribes from $A$, adjusting for individual voters' preferences: $b_i = a_i + v_i$. Therefore, $B$ solves

@@

To fully describe the solution to equation \ref{eqn:solution}, we note the following: for any $a \in \mathbb{R}_+^n$, let $C(a) : {i \in N : a_i > 0}$ denote the set of individuals who receive a bribe from $A$. One can show that there is a bribe schedule $a'$ such that for any $i,j \in C(a)$, $a'_i + v_i = a'_j + v_j$. The intuition is that $A$ has no incentive to make voters differentially bribed, because $B$ will simply ignore the more expensive voters and target the weakest rings in the chain. Following Groseclose and Snyder (1996) we refer to this as a leveling schedule.

Let $U^l(\mathbf{v}, W_B) \subseteq U(\mathbf{v}, W_B)$ denote the set of unbeatable leveling schedules. These are bribe schedules such that $a_i + v_i = a_j + v_j  \equiv t(a)$. The bribe $a_i = t(a) - v_i$ is the sum of two terms. The first is the common "transfer" among all voters in $C(a)$, the second ($-v_i$) is individual specific. The latter term makes voters indifferent between $x$ and $y$ absent any bribe from $B$; the former represents the per capita amount necessary to make $C(a)$, together with any unbribed voters, unaffordable for $B$.

@@

\begin{align*}
    A_1 &: v_{(n+1)/2} < 0\\
    A_2 &: v_1 < 2W_B/(n+1) 
\end{align*}

$A_1$ implies that absent any bribes by $A$, $y$ will defeat $x$. Therefore $A$ must bribe at least one voter. $A_2$ further implies that $A$ must bribe at least a majority of voters, otherwise $B$ will have sufficient resources to bribe $(n+1)/2$ voters and win.

Banks then proceeds to show that there are monotonic bribing schedules contained within the solution for equation \ref{eqn:solution}. For any $a \in \real$ let $k(a) = \lvert C(a) \rvert$. Suppose that $a\in U^l(\mathbf{v}, W_B)$ is such that $v_i \geq v_j$ and $j \in C(a)$ but $i \notin C(a)$. Then, under $A_2$, there exists $a' \in U^l(\mathbf{v}, W_B)$ with $S(a') \leq S(a)$, $k(a') = k(a)$ and $i \in C(a')$ but $j \notin C(a')$ by simply swapping $i$ for $j$.\footnote{Note that since $v_i \geq v_j$, we have that $t(a) - v_i \leq  t(a) - v_j$, i.e. $a'_i \leq a_j$. $A_2$ guarantees that $a'_i$ and $a_j$ are non-negative.}

@@

Therefore, when $A_2$ holds,

@@

subject to the constraint that the induced schedule $a \in U_m^l$. Banks then reformulates this as an unconstrained problem by noting the following. First, if $a(k, t, \mathbf{v})$ is unbeatable, it must be that $k \geq (n + 1)/2$, so by $A_1$ it must be that if $a_i(k, t, \mathbf{v}) = 0$, then $v_i < 0$. Therefore, $B$ receives all non-bribed voters for free. For $a(k, t, \mathbf{v})$ to be unbeatable, then, it must be that $B$ cannot afford the remaining $(n + 1)/2 - (n - k) = k - (n - 1)/2$ voters, or

$$t * (k - (n - 1)/2) \geq W_B$$

@@

Denote the solution to expression \ref{eqn:minimal_expenditure} as $k^*(\mathbf{v}, W_B)$. This solution genenerates a solution to expression \ref{eqn:solution}

@@

Next, suppose that $\Delta(k)$ is increasing in $k$. The following algorithm can then be used to identify $k^*$: if $\Delta((n + 1)/2) \geq 0$, then we know from $]\Delta(k)$ increasing that $A$ is better off by setting $k^*$ to $(n + 1)/2$. If $\Delta((n + 1)/2) < 0$, then we know that $k^*$ must be greater than $(n + 1)/2$, so we next solve for $\Delta((n + 3)/2)$, and so on.

@@

\begin{align}
    \Delta(k)  & = \left[\frac{(k + 1)W_B}{k + 1 - (n - 1)/2} - \sum_{i \leq k + 1}v_i \right]\\
    \label{eqn:delta_k}
    & = \frac{-W_B (n - 1)}{2(k + 1 - (n - 1)/2)(k - (n - 1)/2)} - v_{k + 1}\\
    \label{eqn:def_delta_k}
    & \equiv T(k, W_B) - v_{k + 1}
\end{align}

@@

Generating an explicit characterization of $k$ is not straightforward. Instead, Banks opts to show the conditions which need to hold in order for the coalition to be universal or minimal. Using equation \ref{eqn:optimal_k} and substituting in equation \ref{eqn:delta_k} we have the following.

@@

From equation \ref{eqn:optimal_k} we infer that $\Delta(k - 1, \mathbf{v}', W_B) < 0)$, which from equations \ref{eqn:delta_k} and \ref{eqn:def_delta_k} is equivalent to $v_k' > T(k' - 1, W_B)$. Now suppose that the preference profile changes from $\mathbf{v}'$ to $\mathbf{v}$, and $v_{k'}$ is such that $v_{k'} \geq v'_k$. Then, $v_{k'} > T(k' - 1, W_B)$, and hence $\Delta(k' - 1, \mathbf{v}, W_B) < 0$. But then from equation \ref{eqn:optimal_k} it must be the case that $k^*(\mathbf{v}, W_B) \geq k'$. Therefore, the following holds:

>>>>

Two states of the world: policies $x$ and $y$. Party $A$ and $B$ prefer $x$ and $y$ respectively. We can make this more explicit with
\begin{align*}
    U_A(x) > U_A(y)\\
    U_B(y) > U_B(x)
\end{align*}
For voters, $v_i > 0$ means voter $i$ prefers $x$ to $y$. All $i \in N$ vote for $x$ or $y$. Simple majority rule determines which policy gets implemented. For each voter, each party sets a bribe schedule
\begin{align*}
    a \in (A1, ..., An) \in \mathbb{R}^n_+\\
    b \in (b_1, ..., b_n) \in \mathbb{R}^n_+
\end{align*}
Solving through backward induction, given bribe schedules $(a,b)$, voter $i$ prefers to vote for $x$ if $a_i + v_i > b_i$ and for $y$ otherwise. Since indifferent voters choose $y$, party $B$ needs to only match bribes from $A$, adjusting for individual voters' preferences: $b_i = a_i + v_i$. Therefore, $B$ solves

@@

To fully describe the solution to equation \ref{eqn:solution}, we note the following: for any $a \in \mathbb{R}_+^n$, let $C(a) = \{i \in N : a_i > 0\}$ denote the set of individuals who receive a bribe from $A$. One can show that there is a bribe schedule $a'$ such that for any $i,j \in C(a)$, $a'_i + v_i = a'_j + v_j$. The intuition is that $A$ has no incentive to make voters differentially bribed, because $B$ will simply ignore the more expensive voters and target the weakest rings in the chain. Following Groseclose and Snyder (1996) we refer to this as a leveling schedule.

Let $U^l(\mathbf{v}, W_B) \subseteq U(\mathbf{v}, W_B)$ denote the set of unbeatable leveling schedules. These are bribe schedules such that $a_i + v_i = Aj + v_j  \equiv t(a)$. The bribe $a_i = t(a) - v_i$ is the sum of two terms. The first is the common "transfer" among all voters in $C(a)$, the second ($-v_i$) is individual specific. The latter term makes voters indifferent between $x$ and $y$ absent any bribe from $B$; the former represents the per capita amount necessary to make $C(a)$, together with any unbribed voters, unaffordable for $B$.

@@

\begin{align*}
    A1 &: v_{(n+1)/2} < 0\\
    A2 &: v_1 < 2W_B/(n+1) 
\end{align*}

$A1$ implies that absent any bribes by $A$, $y$ will defeat $x$. Therefore $A$ must bribe at least one voter. $A2$ implies that $A$ must bribe at least a majority of voters, otherwise $B$ will have sufficient resources to bribe $(n+1)/2$ voters and win. $A2$ also implies that for all $a \in U^l(\mathbf{v}, W_B)$, it must be that $t(a) \geq 2W_B/(n + 1)$, otherwise $B$ can bribe a majority from $C(a)$ itself and win the vote.

Banks then proceeds to show that there are monotonic bribing schedules contained within the solution for equation \ref{eqn:solution}. For any $a \in \real$ let $k(a) = \lvert C(a) \rvert$. Suppose that $a\in U^l(\mathbf{v}, W_B)$ is such that $v_i \geq v_j$ and $j \in C(a)$ but $i \notin C(a)$. Then, under $A2$, there exists $a' \in U^l(\mathbf{v}, W_B)$ with $S(a') \leq S(a)$, $k(a') = k(a)$ and $i \in C(a')$ but $j \notin C(a')$ by simply swapping $i$ for $j$: $a'_i = t(a) - v_i$, $a'_j = 0$ and $a'_m = a_m$ for all $m \notin i, j$.\footnote{Note that since $v_i \geq v_j$, we have that $t(a) - v_i \leq  t(a) - v_j$, i.e. $a'_i \leq a_j$. $A2$ guarantees that $a'_i$ and $a_j$ are non-negative.}

@@

Therefore, when $A2$ holds,

@@

subject to the constraint that the induced schedule $a \in U_m^l$. Banks then reformulates this as an unconstrained problem by noting the following. First, if $a(k, t, \mathbf{v})$ is unbeatable, it must be that $k \geq (n + 1)/2$, so by $A1$ it must be that if $a_i(k, t, \mathbf{v}) = 0$, then $v_i < 0$. Therefore, $B$ receives all non-bribed voters for free. For $a(k, t, \mathbf{v})$ to be unbeatable, then, it must be that $B$ cannot afford the remaining $(n + 1)/2 - (n - k) = k - (n - 1)/2$ voters, or

$$t \cdot (k - (n - 1)/2) \geq W_B$$

@@

Denote the solution to expression \ref{eqn:minimal_expenditure} as $k^*(\mathbf{v}, W_B)$. This solution implicitly generates a solution to expression \ref{eqn:solution} through expression \ref{eqn:total_expenditure} and the induced bribe schedule above.

@@

Next, suppose that $\Delta(k)$ is increasing in $k$. The following algorithm can then be used to identify $k^*$: if $\Delta((n + 1)/2) \geq 0$, then we know from $\Delta(k)$ increasing that $A$ is better off by setting $k^*$ to $(n + 1)/2$. If $\Delta((n + 1)/2) < 0$, then we know that $k^*$ must be greater than $(n + 1)/2$, so we next solve for $\Delta((n + 3)/2)$, and so on.

@@

\begin{align}
    \Delta(k)  & = \left[\frac{(k + 1)W_B}{k + 1 - (n - 1)/2} - \sum_{i \leq k + 1}v_i \right]\\
    \label{eqn:deltAk}
    & = \frac{-W_B (n - 1)}{2(k + 1 - (n - 1)/2)(k - (n - 1)/2)} - v_{k + 1}\\
    \label{eqn:def_deltAk}
    & \equiv T(k, W_B) - v_{k + 1}
\end{align}

@@

Generating an explicit characterization of $k$ is not straightforward. Instead, Banks opts to show the conditions which need to hold in order for the coalition to be universal or minimal. Using equation \ref{eqn:optimal_k} and substituting in equation \ref{eqn:deltAk} we have the following.

@@

From equation \ref{eqn:optimal_k} we infer that $\Delta(k - 1, \mathbf{v}', W_B) < 0)$, which from equations \ref{eqn:deltAk} and \ref{eqn:def_deltAk} is equivalent to $v_k' > T(k' - 1, W_B)$. Now suppose that the preference profile changes from $\mathbf{v}'$ to $\mathbf{v}$, and $v_{k'}$ is such that $v_{k'} \geq v'_k$. Then, $v_{k'} > T(k' - 1, W_B)$, and hence $\Delta(k' - 1, \mathbf{v}, W_B) < 0$. But then from equation \ref{eqn:optimal_k} it must be the case that $k^*(\mathbf{v}, W_B) \geq k'$. Therefore, the following holds:
